---
title: "QTM 150 Final Project"
author: "Anna Choi"
output: html_document
---

## 0. Setup
```{r}
library(tidyverse)
setwd("~/Desktop/QTM 150")
cacd2016 <- read_csv("cacd2016.csv")

# Data on Federal cases assigned in 2016 to judges in the Central District of California
cacd2016
```

## 1. Distribution of cases in cacd2016
```{r}
# A histogram of the number of cases that the judge was assigned to in 2016 in cacd2016
ggplot(cacd2016, aes(N.Cases.2016)) +
  geom_histogram()

# A density plot of the number of cases that the judge was assigned to in 2016 in cacd2016
ggplot(cacd2016, aes(N.Cases.2016)) +
  geom_density()
```

## 2. Distribution of cases under simulated random assignment
### Simulate a random assignment process for the number of judges and cases in cacd2016
```{r}
# The total number of cases in cacd2016
n.cases <- sum(cacd2016$N.Cases.2016)
n.cases

# The total number of judges in cacd2016
n.judges <- cacd2016 %>% nrow()
n.judges

# A simulated random assignment process for n.judges and n.cases
simulated <- rep(0,91)
for(i in 1:10501){
  ith_case <- sample(1:91, 1)
  simulated[ith_case] <- simulated[ith_case] + 1
  }
simulated

# A histogram of the distribution of cases that resulted from the simulated random assignment process
ggplot(cacd2016, aes(simulated)) +
  geom_histogram() +
  labs(x = "N.Cases")

# A density plot of the distribution of cases that resulted from the simulated random assignment process
ggplot(cacd2016, aes(simulated)) +
  geom_density() +
  labs(x = "N.Cases")
```

## 3. Dealing with outliers
### i) Visiting hypothesis
```{r}
# Create and add a new variable fewer_than_75_cases to cacd2016 where fewer_than_75_cases represents whether the number of cases that the judge was assigned to is fewer than 75 or not
cacd2016 <- mutate(cacd2016, fewer_than_75_cases = N.Cases.2016 < 75)

# Make variable Birth.State binary by looking at CA vs. Non-CA
birth_state_binary <- cacd2016 %>%
  mutate(from_CA = Birth.State == "CA")

# A bar plot representing the number of judges in cacd2016 with fewer than 75 cases vs. the number of judges with at least 75 cases
ggplot(birth_state_binary, aes(fewer_than_75_cases)) + 
  geom_bar(aes(fill = from_CA), position = "dodge") +
  labs(y = "n.judges")
```
```{r}
birth_state_binary$fewer_than_75_cases[birth_state_binary$fewer_than_75_cases == "TRUE"] <- "Fewer than 75"
birth_state_binary$fewer_than_75_cases[birth_state_binary$fewer_than_75_cases == "FALSE"] <- "More than 75"

birth_state_binary$fewer_than_75_cases <- factor(birth_state_binary$fewer_than_75_cases, levels = c("More than 75", "Fewer than 75"))

ggplot(birth_state_binary, aes(fewer_than_75_cases)) + 
  geom_bar(aes(fill = from_CA), position = "dodge") +
  labs(x = "Number of cases in 2016", y = "Number of Judges") +
  guides(fill = guide_legend(title = "Born in California"))
```
```{r}
birth_state_binary <- birth_state_binary %>%
  group_by(fewer_than_75_cases, from_CA) %>%
  summarize(n = n()) %>%
  mutate(percentage = (n/sum(n))*100)
birth_state_binary
```
```{r}
ggplot(birth_state_binary, aes(fewer_than_75_cases, percentage)) + 
  geom_bar(aes(fill = from_CA), stat = "identity", position = "dodge") +
  labs(x = "Number of cases in 2016", y = "Percentage of Judges") +
  guides(fill = guide_legend(title = "Born in California"))
```

### ii) Removing visitors
```{r}
# Create a new dataframe called local_cacd2016 containing only the judges in cacd2016 who where assigned more than 75 cases in 2016
local_cacd2016 <- filter(cacd2016, fewer_than_75_cases == "FALSE")

# A density plot of N.Cases.2016 in local_cacd2016
ggplot(local_cacd2016, aes(N.Cases.2016)) +
  geom_density()
```
```{r}
# The total number of cases in local_cacd2016
local_n.judges.cases <- sum(local_cacd2016$N.Cases.2016)
local_n.judges.cases

# The total number of judges in local_cacd2016
local_n.judges.judges <- local_cacd2016 %>% nrow()
local_n.judges.judges

# A simulated random assignment process for local_n.judges.judges and local_n.judges.cases
local_simulated <- rep(0,27)
for(i in 1:10331){
  ith_case <- sample(1:27, 1)
  local_simulated[ith_case] <- local_simulated[ith_case] + 1
  }
local_simulated

# A density plot of the distribution of cases that resulted from the new simulated random assignment process
ggplot(local_cacd2016, aes(local_simulated)) +
  geom_density() +
  labs(x = "N.Cases")
```

## 4. Comparisons across salient groups (after removing visitors)
### i) Raw comparisons
```{r}
# The average number of cases assigned to Female and Male judges in local_cacd2016
local_cacd2016 %>%
  group_by(Gender) %>%
  summarize(avg_n.cases = mean(N.Cases.2016)) %>%
  select(Gender, avg_n.cases)
```
```{r}
# Make variable Race.or.Ethnicity binary by looking at White vs. Non-White
race_or_ethnicity_binary <- local_cacd2016 %>%
  mutate(White = Race.or.Ethnicity == "White")

# The average number of cases assigned to White and Non-White judges in local_cacd2016
race_or_ethnicity_binary %>%
  group_by(White) %>%
  summarize(avg_n.cases = mean(N.Cases.2016)) %>%
  select(White, avg_n.cases)
```
```{r}
# The average number of cases assigned across all races and ethnicities (Race.or.Ethnicity) in local_cacd2016
local_cacd2016 %>%
  group_by(Race.or.Ethnicity) %>%
  summarize(avg_n.cases = mean(N.Cases.2016)) %>%
  select(Race.or.Ethnicity, avg_n.cases)
```

### ii) Statistical test for race differences
```{r}
# The number of judges and the number of assigned cases of race groups in local_cacd2016
race_comp_test <- race_or_ethnicity_binary %>%
  group_by(White) %>%
  summarize(n.cases = sum(N.Cases.2016), n.judges = n()) %>%
  select(White, n.cases, n.judges)
race_comp_test
```
```{r}
binom_test <- function(ds){
  sX1 <- ds %>% slice(1) %>% pull(n.cases)
  nX1 <- ds %>% slice(1) %>% pull(n.judges)
  m <- sum(ds$n.cases)
  n <- sum(ds$n.judges)
  
  Z <- (sX1 - m*(nX1/n))/((m*(nX1/n)*(1-(nX1/n)))^(1/2))
  p_val <- 2*pnorm(-abs(Z))
  
  return(p_val)
}

# Run binom_test with race_comp_test
binom_test(race_comp_test)
```

## 5. Multiple comparisons
### i) Race differences
```{r}
# Make variable Race.or.Ethnicity binary by looking at African American or Hispanics vs. the rest
race_or_ethnicity_binary <- local_cacd2016 %>%
  mutate(African.American.or.Hispanic = Race.or.Ethnicity %in% c("African American", "Hispanic"))

# The number of judges and the number of assigned cases of race groups in local_cacd2016
race_comp_test_2 <- race_or_ethnicity_binary %>%
  group_by(African.American.or.Hispanic) %>%
  summarize(n.cases = sum(N.Cases.2016), n.judges = n()) %>%
  select(African.American.or.Hispanic, n.cases, n.judges)
race_comp_test_2

# Run binom_test with race_comp_test_2
binom_test(race_comp_test_2)
```

### ii) Robust Checks
```{r}
# Create and add a new variable born_before_1945 to local_cacd2016 where born_before_1945 represents whether the judge was born before 1945 or not
local_cacd2016 <- mutate(local_cacd2016, born_before_1945 = Birth.Year < 1945)

# Make variable Birth.Year binary by looking at judges born before 1945 vs. after 1945
age_binary <- local_cacd2016 %>%
  mutate(Born.Before.1945 = Birth.Year < 1945)

# The number of judges and the number of assigned cases of age groups in local_cacd2016
age_comp_test <- age_binary %>%
  group_by(Born.Before.1945) %>%
  summarize(n.cases = sum(N.Cases.2016), n.judges = n()) %>%
  select(Born.Before.1945, n.cases, n.judges)
age_comp_test

# Run binom_test with age_comp_test
binom_test(age_comp_test)
```

```{r}
# Create and add a new variable born_before_1950 to local_cacd2016 where born_before_1950 represents whether the judge was born before 1950 or not
local_cacd2016 <- mutate(local_cacd2016, born_before_1950 = Birth.Year < 1950)

# Make variable Birth.Year binary by looking at judges born before 1950 vs. after 1950
age_binary <- local_cacd2016 %>%
  mutate(Born.Before.1950 = Birth.Year < 1950)

# The number of judges and the number of assigned cases of age groups in local_cacd2016
age_comp_test_2 <- age_binary %>%
  group_by(Born.Before.1950) %>%
  summarize(n.cases = sum(N.Cases.2016), n.judges = n()) %>%
  select(Born.Before.1950, n.cases, n.judges)
age_comp_test_2

# Run binom_test with age_comp_test_2
binom_test(age_comp_test_2)
```

```{r}
# The minimum Birth.Year in local_cacd2016
min(local_cacd2016$Birth.Year)

# The maximum Birth.Year in local_cacd2016
max(local_cacd2016$Birth.Year)

# The collection of all possible values in Birth.Year in ascending order, excluding repetitions
sort(unique(local_cacd2016$Birth.Year))
```

```{r}
birth.y_tests <- data.frame()

for(i in sort(unique(local_cacd2016$Birth.Year))){
  local_cacd2016 <- mutate(local_cacd2016, born_before_i = Birth.Year < i)
  age_binary <- local_cacd2016 %>%
    mutate(Born.Before.i = Birth.Year < i)
  age_binary$Born.Before.i[age_binary$Born.Before.i == "TRUE"] <- "Old"
  age_binary$Born.Before.i[age_binary$Born.Before.i == "FALSE"] <- "Young"
  age_comp_test_i <- age_binary %>%
    group_by(Born.Before.i) %>%
    summarize(n.cases = sum(N.Cases.2016), n.judges = n()) %>%
    select(Born.Before.i, n.cases, n.judges)
  binom_test(age_comp_test_i)
  output = c(i, binom_test(age_comp_test_i))
  birth.y_tests = rbind(birth.y_tests, output)
  print(i)
  print(binom_test(age_comp_test_i))
  print(age_comp_test_i)
}

# Name the columns birth.y indicating the year threshold and test_results indicating the result of binom_test for that year
colnames(birth.y_tests) <- c("birth.y", "test_results")
birth.y_tests <- birth.y_tests[order(birth.y_tests$birth.y),]
birth.y_tests
```
```{r}
# A histogram of test_results in birth.y_tests
ggplot(birth.y_tests, aes(test_results)) +
  geom_histogram()
```
```{r}
# Create and add a new variable below_0.05 to birth.y_tests where below_0.05 represents whether the binom_test result is smaller than 0.05 or not
birth.y_tests <- mutate(birth.y_tests, below_0.05 = test_results < 0.05)

# Create and add a new variable below_0.01 to birth.y_tests where below_0.01 represents whether the binom_test result is smaller than 0.01 or not
birth.y_tests <- mutate(birth.y_tests, below_0.01 = test_results < 0.01)

# A scatter plot based on birth.y_tests with birth.y on the x-axis and test_results on the y-axis
ggplot(birth.y_tests, aes(birth.y, test_results)) +
  geom_point(aes(color = below_0.05, shape = below_0.01))
```
```{r}
# A histogram of the percentage of test_results in birth.y_tests
ggplot(birth.y_tests, aes(test_results)) +
  geom_histogram(aes(y = stat(count) / sum(count))) +
  labs(x = "P-values", y = "Percentage") + 
  geom_vline(xintercept = 0.05, linetype = "dashed", color = "red")
```
```{r}
# Make the variable birth.y represent an age threshold in 2016
birth.y_tests <- mutate(birth.y_tests, birth.y = 2016 - birth.y)

# A scatter plot based on birth.y_tests with Age threshold on the x-axis and P-values on the y-axis
ggplot(birth.y_tests, aes(birth.y, test_results)) +
  geom_point(aes(color = below_0.05, shape = below_0.01)) +
  labs(x = "Age threshold (young vs. old)", y = "P-values", color = "Below 0.05", shape = "Below 0.01")
```
