---
title: "HW 2: qplot and data import"
author: "Anna Choi"
output: html_document
---

## 0. Setup
```{r}
library('tidyverse')
```

## 1. Population density and poverty in the Midwest
```{r}
midwest
```

### 1-a. Scatter plot illustrating the relationship between the population density of a county and its percent of population below poverty

```{r}
qplot(x = popdensity, y = percbelowpoverty, data = midwest)
```

### 1-b. Scatter plots illustrating the relationship between the population density of a county and its percent of population below poverty broken down by whether the county is in a metropolitan area or not

```{r}
qplot(popdensity, percbelowpoverty, data = midwest, facets = ~inmetro)
```

## 2. Historical trends for obesity and diabetes in the US

### 2-a. Import "obes_diab_trend.csv" data set

```{r}
setwd("~/Desktop/QTM 150")
diabobestrend <- read_csv("obes_diab_trend.csv")
diabobestrend
```

### 2-b. The historical trend of the prevalence of obesity and diabetes

```{r}
qplot(x = year,
      y = prevalence,
      data = diabobestrend,
      geom = 'line',
      color = condition)
```

## 3. Challenge question (extra credit)

### 3-a. Scatter plot illustrating the relationship between the population density of a county and its percent of population below poverty for counties with population density no larger than 10,000 only

```{r}
library(dplyr)
midwest_filter <- filter(midwest, popdensity < 10000)
qplot(popdensity,
      percbelowpoverty,
      data = midwest_filter,
      xlab = "Population Density",
      ylab = "Percent below Poverty")
```
