---
title: "HW 3: dplyr"
author: "Anna Choi"
output: html_document
---

## 0. Setup
```{r}
library(babynames)
library(tidyverse)
library(gapminder)
```


## 1. Dplyr
```{r}
babynames
```

### 1-a. For the year 2017, find the most common name among females.
```{r}
babynames %>%
  filter(year == 2017, sex == "F") %>%
  arrange(desc(n))
```
#### Emma is the most common name among females in 2017.

### 1-b. For the year 1982, find the number of males.
```{r}
babynames %>%
  filter(year == 1982) %>%
  mutate(counter = 1) %>%
  group_by(sex) %>%
  summarize(n = sum(counter))
```
#### There were 7364 males in 1982.

### 1-c. For each year, find the number of distinct male names and the number of distinct female names given in that year.
```{r}
babynames %>%
  mutate(counter = 1) %>%
  group_by(year, sex) %>%
  summarize(n = sum(counter))
```


## 2. Dplyr + qplot

### 2-a. Find the four most given names in 2017 (across both sexes). Store these four names into a character vector and call that vectortop_names_2017.
```{r}
babynames %>%
  filter(year == 2017) %>%
  arrange(desc(n))
```
#### Emma, Liam, Olivia, and Noah are the four most given names in 2017.

```{r}
vectortop_names_2017 <- babynames %>%
  filter(name %in% c("Emma", "Liam", "Olivia", "Noah"))
```

### 2-b. Use a line graph to display the evolution through time of the proportion of all given names that correspond to one of these four names (one line per name).
```{r}
ggplot(vectortop_names_2017, aes(x = year, y = prop)) +
  geom_line(aes(color = name))
```

### 2-c. The sudden drops in the previous graph come from the fact that it depicts two observations for each year. For each name, the graph represents the number of females and males who were given the name.
```{r}
vectortop_names_2017 <- babynames %>%
  filter(name %in% c("Emma", "Liam", "Olivia", "Noah"), prop > 0.0002)

ggplot(vectortop_names_2017, aes(x = year, y = prop)) +
  geom_line(aes(color = name))
```
