---
title: 'HW 4: ggplot'
author: "Anna Choi"
output: html_document
---

## 0. Setup
```{r}
library(tidyverse)
setwd("~/Desktop/QTM 150")
newgss <- read_csv("newgss.csv")
```


## 1. GSS
```{r}
newgss
```

### 1-a. Using newgss, create a barplot of degree by sex.
```{r}
ggplot(newgss, aes(degree)) + 
  geom_bar(aes(fill = sex))
```

### 1-b. Create another barplot of degree by sex, dividing the graph into region.
```{r}
ggplot(newgss, aes(degree)) + 
  geom_bar(aes(fill = sex)) +
  facet_wrap(~region)
```

### 1-c. Compute the average of income1 for each combination of year and sex.
```{r}
newgss_sum <- newgss %>% group_by(year, sex) %>% summarize(avg_inc = mean(income1, na.rm = TRUE))
newgss_sum
```

### 1-d. Based on the averages you computed in “1-c”, create a line graph showing the evolution of average income through time by sex (set the color of each line to the variable sex).
```{r}
ggplot(newgss_sum, aes(year, avg_inc)) + 
  geom_line(aes(color = sex))
```


## 2. Txhousing
```{r}
txhousing
```

### 2-a. Filter the txhousing data frame.
```{r}
txhousing_filt <- filter(txhousing, year == 2014, city %in% c("Austin", "Collin County", "Dallas", "Denton County", "Fort Bend", "Fort Worth", "Houston", "Montgomery County", "San Antonio"))
txhousing_filt
```

### 2-b. Based on the data frame you created in “2-a”, create a graph showing the evolution of median through the 12 months of the year.
```{r}
ggplot(txhousing_filt, aes(month, median)) + 
  geom_point(aes(color = city), size = 0.5) +
  geom_line(aes(color = city), alpha = 0.5)
```


## 3. Challenge question (Extra Credit)
```{r}
newgss
```

### 3-a. Based on newgss, produce a data frame called married_happy.
```{r}
married_happy <- filter(newgss, !happy %in% c("DK", NA), !marital %in% c(NA))
married_happy <- married_happy %>%
  group_by(marital, happy) %>%
  summarize(n = n()) %>%
  mutate(prop = n/sum(n)) %>%
  select(marital, happy, prop)
married_happy
```

### 3-b. Provide a bar plot of married_happy.
```{r}
ggplot(married_happy, aes(marital, prop)) + 
  geom_bar(aes(fill = happy), stat = "identity", position = "dodge") +
  labs(y = "Proportion", x = "Marital Status") +
  guides(fill = guide_legend(title = "Happiness")) +
  theme(axis.text.x = element_text(size = 8))
```
